---
title: "Watching R progress with fernsicht"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Watching R progress with fernsicht}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE       # All chunks are non-evaluating: they touch the
                     # network and spawn the bridge subprocess.
)
```

`fernsicht` lets you watch any long-running R loop from a browser, on
any device, without an account or installed app. Progress flows
peer-to-peer over WebRTC, so the data stays between you and your
viewers — the central server only mediates the handshake.

This vignette walks through the three API layers and when to reach
for each.

## Install

```{r}
remotes::install_github("MuteJester/Fernsicht", subdir = "publishers/r")
```

The R package is small (~1,200 lines of R, no compiled code). On
first use it downloads a ~10 MB Go binary (`fernsicht-bridge`) from
GitHub releases, caches it under `tools::R_user_dir("fernsicht",
"cache")`, and verifies its SHA256 against a table bundled in the
package.

## Layer 1 — `blick()`

The headline. A drop-in replacement for `lapply()` that streams
progress while it runs.

```{r}
library(fernsicht)

results <- blick(1:100, function(i) {
  Sys.sleep(0.05)
  i * 2
})
#> Connecting to https://signal.fernsicht.space ...
#> Connected. Viewer: https://app.fernsicht.space/#room=abc12345
```

Open the printed URL on your phone, send it to a colleague, project
it on a wall — they'll see the bar fill in live as your loop runs.

`blick()` preserves names if `items` has them (matching `lapply()`):

```{r}
named <- c(a = 1, b = 2, c = 3)
out <- blick(named, function(x) x + 100)
#>   a   b   c
#> 101 102 103
```

A second `blick()` in the same R session **reuses the same viewer
URL** — no new message, no new link to share. The session lives for
the rest of the R process unless you `close_session()` it.

## Layer 2 — `with_session()` for manual loops

Sometimes a `for` loop is the right shape. `with_session()` scopes a
task around an expression so bare `tick()` calls inside know which
task to update.

```{r}
with_session({
  for (i in seq_len(100)) {
    do_one_step(i)
    tick(value = i / 100, n = i, total = 100)
  }
}, label = "Bootstrap iteration")
```

Nested `with_session()` works — the inner scope replaces the outer
on the wire, and on exit the outer task resumes so subsequent ticks
land on it again.

## Layer 3 — manual lifecycle

For multi-phase work (e.g., training epochs containing batches):

```{r}
sess <- session(label = "Training")

for (epoch in 1:10) {
  start_task(sess, "epoch", paste("Epoch", epoch))
  for (batch in 1:100) {
    train_batch()
    tick(sess, "epoch",
         value = batch / 100, n = batch, total = 100)
  }
  end_task(sess, "epoch")
}

close_session(sess)   # optional — exits anyway when R exits
```

`session()` returns the per-process **ambient session**. Calling it a
second time returns the same one — no new bridge subprocess, no new
viewer URL.

## Accessors

```{r}
viewer_url()                # ambient session's URL
viewer_url(copy = TRUE)     # also copy to clipboard (needs `clipr`)
viewers()                   # current roster, e.g. c("vega", "orion")
diagnostics()               # snapshot suitable for logs / bug reports
```

`diagnostics()` is safe to log: the bridge's sender secret is
**never** included in any public accessor or printable form.

## Non-interactive environments

In `Rscript`, `R CMD BATCH`, or RStudio's Job pane, `message()`
output may not be visible. Use the accessor:

```{r}
sess <- session()
cat("share this URL:", viewer_url(sess), "\n")
```

Or set the env var `FERNSICHT_VIEWER_URL_SINK` (planned in v1.x) to
write the URL to a file on session creation.

## Configuration

Resolution order: function argument > R option > environment variable > default.

| Setting | Argument | Option | Env var | Default |
|---|---|---|---|---|
| Server URL | `server_url=` | `fernsicht.server_url` | `FERNSICHT_SERVER_URL` | `https://signal.fernsicht.space` |
| Join secret | `join_secret=` | `fernsicht.join_secret` | `FERNSICHT_JOIN_SECRET` | (none) |
| Max viewers | `max_viewers=` | `fernsicht.max_viewers` | `FERNSICHT_MAX_VIEWERS` | `8` |
| Bridge binary | `bridge_path=` | `fernsicht.bridge_path` | `FERNSICHT_BRIDGE_PATH` | (cache) |

`FERNSICHT_BRIDGE_PATH` is the development override — useful when
hacking on the bridge itself.

## Errors

Conditions are typed for `tryCatch()`:

```{r}
tryCatch(
  blick(1:10, my_fn),
  fernsicht_session_error    = function(e) message("session failed: ", e$message),
  fernsicht_session_expired  = function(e) message("session expired (12h limit)"),
  fernsicht_signaling_error  = function(e) message("can't reach signal server"),
  fernsicht_internal_error   = function(e) message("bridge crashed")
)
```

Sessions cap at 12 hours. The SDK warns once at ~80% elapsed so
week-long runs can plan a restart.

## Limits

- **One session per R process.** Open multiple R processes for
  multiple sessions. `furrr` / `future` parallel support is on the
  roadmap.
- **12-hour session lifetime.** Restart manually for longer jobs;
  token refresh is planned but requires bridge work.
- **No telemetry.** The package only contacts the configured
  signaling server and GitHub releases.

## Going further

- [`?blick`](../html/blick.html), [`?with_session`](../html/with_session.html),
  [`?session`](../html/session.html) for full reference docs.
- The bridge protocol spec
  ([BRIDGE_PROTOCOL.md](https://github.com/MuteJester/Fernsicht/blob/main/BRIDGE_PROTOCOL.md))
  documents the wire format if you want to write an SDK for another language.
